---
title: 'Homework 3: Sims & Viz'
author: "Mickayla Johnston"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
# Libraries
library(dplyr)
library(ggplot2)
library(readr)
library(forcats)
library(TSP)
```


#### 1. Sample Properties  
##### Consider the following vasopressin levels in voles.

```{r}
vole_vaso <- c(98,96,94,88,86,82,77,74,70,60,
                59,52,50,47,40,35,29,13,6,5)
```
  

##### 1a. Say “Vole vasopressin” 10 times as fast as you can. How many times did you trip up?  
None!! :-)  

##### 1b. What is the mean, median, sd, and interquartile range of the sample?  
```{r}
mean(vole_vaso)
```
The mean is 58.05    

```{r}
median(vole_vaso)
```
The median is 59.5    

```{r}
sd(vole_vaso)
```
The SD is 29.75244  

```{r}
IQR(vole_vaso)
```
The IQR is 44.25  

##### 1c. What is the standard error of the mean (do this with a formula!)?  

SE of the mean = sd of your sample / the square root of your sample size
```{r}
vole_se_mean <- sd(vole_vaso) / sqrt(length(vole_vaso))
vole_se_mean
```
SE of the mean = 6.652849  

##### 1d. What does the standard error of the mean tell you about our estimate of the mean values of the population of vole vassopressin?  
The SE tells us how precise our samples are, i.e., how close to the actual population mean our sample mean is.  

#### 2. Sample Size for upper quartiles  
We can get the upper quartile value of vole vassopressin with
```{r}
quantile(vole_vaso, probs = 0.75)
```
75% = 83

##### Let’s assume the sample is representative of the popultion.  
##### 2a. Use sample() to get just one resample with a sample size of 10. What is its upper quartile?  

```{r}
# Create sample size object
samp_size <- 10

# Set seed for reproducibility
set.seed(123)

# Run 1 resample using sample size object to get the upper quantile
sample(vole_vaso,
       size = samp_size,
       replace = TRUE) %>%
  quantile(prob = 0.75)
```
The upper quartile of this resample is 59.75  

##### 2b. Build an initial data frame for simulations with the sample sizes 5 through 20.  
```{r}
# Set seed for reproducibility
set.seed(123)

# Create data frame
vole_upperq_sims <- data.frame(samp_sizes = 5:20) %>%

# 2c. Use this data frame to get simulated upper quartiles for each sample size using 1,000 simulations

  rowwise(samp_sizes) %>%
  # replicate 
  summarize(boot_upperq = replicate(1000,
                                  # bootstrap draws
                                  sample(vole_vaso,
                                         size = samp_sizes,
                                         replace = TRUE) %>%
                                    # & get an IQR
                                    quantile(prob = 0.75)))

# View the output
vole_upperq_sims
```

##### 2d. With a ggplot, make a guesstimate as to the best sample size for estimating the upper quartile of the population. Use whatever geom you feel makes things most easy to see.  
E.C. Add a red dashed line using geom_vline() or geom_hline() to show where that should be, perhaps.  

```{r}
# Create the plot
iqr_samp_size_plot <- ggplot(data = vole_upperq_sims,
       mapping = aes(x = samp_sizes,
                     y = boot_upperq)) +
  # Add geom_point
  geom_point() +
  
  # Change x-axis scale to make plot more clear
  scale_x_continuous(breaks = round(seq(min(vole_upperq_sims$samp_sizes), max(vole_upperq_sims$samp_sizes), by = 1),1))

# View plot
iqr_samp_size_plot
```
  
The best sample size looks to be around 11, since the upper quantile doesn't seem to vary much from sample size of 11 and up

```{r}
# Add vline at this sample size:
iqr_samp_size_plot + geom_vline(xintercept = 11, color = "red",
                                linetype = "dashed")
```
  

##### 2e. Plot the SE of the estimate of the upper quantile by sample size. Again, what it the best way to see this? Does it level off? Is there a level you feel acceptable? Justify your answer. Does this match with what you put in 2d?
```{r}
# Get SE of the upper quantiles
boot_se_upperq <- vole_upperq_sims %>%
  group_by(samp_sizes) %>%
  summarize(se_upperq = sd(boot_upperq))

# Create the plot
se_upperq_plot <- ggplot(data = boot_se_upperq,
                         mapping = aes(x = samp_sizes,
                                       y = se_upperq)) +
  geom_line() +
  # Change x-axis tick marks to make clearer
  scale_x_continuous(breaks = round(seq(min(boot_se_upperq$samp_sizes), max(boot_se_upperq$samp_sizes), by = 1),1))

# View it
se_upperq_plot
```

This plot doesn't seem to level off as much as the previous one. Even at sample size = 11, the SE still seems to continue to decrease. From this plot, a sample size of around 18 seems better, as the SE sort of levels out here.

#### 3. Ggplot  
##### 3a. Some setup. Run the code below. For extra credit, look up the packages and functions used and explain what is going on here. But, that’s EC.  


(Libraries loaded above)  

###### Set theme
```{r}
theme_set(theme_bw(base_size=12))
```

###### Read in data
```{r}
ice <- read_csv("http://biol607.github.io/homework/data/NH_seaice_extent_monthly_1978_2016.csv") %>%
  mutate(Month_Name = factor(Month_Name),
         Month_Name = fct_reorder(Month_Name, Month))
```

##### 3b. Make a boxplot showing the variability in sea ice extent every month.
```{r}

ggplot(data = ice, 
       mapping = aes(x = Month_Name,
                     y = Extent)) +
  geom_boxplot()
```


##### 3c. Use dplyr to get the annual minimum sea ice extent. Plot minimum ice by year. What do you observe?

```{r}
# Calculate annual min ice extent
ice <- ice %>%
  group_by(Year) %>%
  mutate(annual_min_ext = min(Extent)) %>%
  ungroup()

# Check ice for new column
head(ice)

# Plot it by year
ggplot(data = ice,
       mapping = aes(x = Year,
                     y = annual_min_ext)) +
  geom_line()
```
  
Right before 1980 (perhaps 1979), the annual minimum extent drops dramatically. It fluctuates each year after that, hovering around 7, and then drops again around 2002  

##### 3d. One thing that’s really cool about faceting is that you can use cut_*() functions on continuos variables to make facets by groups of continuous variables.  
To see what I mean, try cut_interval(1:10, n = 5) See how it makes five bins of even width? We use cut_interval() or other cut functions with faceting like so facet_wrap(~cut_interval(some_variable)).  
With the original data, plot sea ice by year, with different lines (oh! What geom will you need for that?) for different months. Then, use facet_wrap and cut_interval(Month, n=4) to split the plot into seasons.

###### Try out cut_interval
```{r}
cut_interval(1:10, n = 5)
```

###### Plot of sea ice extent by year, colored by month
```{r}
ext_by_year_plot <- ggplot(data = ice, 
       mapping = aes(x = Year,
                     y = Extent,
                     color = Month_Name)) +
  geom_line()

# View it
ext_by_year_plot
```

###### Add facets by season cut_interval(Month, n = 4)

```{r}
ext_by_year_plot + facet_wrap(~cut_interval(Month, n = 4))
```

##### 3e. Last, make a line plot of sea ice by month with different lines as different years.   
###### Gussy it up with colors by year, a different theme, critical values, and whatever other annotations, changes to axes, etc., you think best show the story of this data. For ideas, see the lab, and look at various palettes around. Extra credit for using colorfindr to make a palette.

```{r}
ext_by_month_plot <- ggplot(data = ice, 
                            mapping = aes(x = Month_Name,
                                          y = Extent,
                                          color = Year)) +
  geom_line(mapping = 
              aes(group = Year))

# View it
ext_by_month_plot
```

```{r}
# Install and load colorfindr
#install.packages("colorfindr")
library(colorfindr)

# Plot (5000 randomly selected pixels) of aurora borealis
get_colors("https://upload.wikimedia.org/wikipedia/commons/a/aa/Polarlicht_2.jpg") %>% 
  plot_colors_3d(sample_size = 5000, marker_size = 2.5, color_space = "RGB")

# Set seed
set.seed(123)

# Get colors and create a palette with n = 5 
my_palette <- get_colors("https://upload.wikimedia.org/wikipedia/commons/a/aa/Polarlicht_2.jpg") %>% 
  make_palette(n = 50)

# Sort colors
rgb <- col2rgb(my_palette)
tsp <- as.TSP(dist(t(rgb)))
sol <- solve_TSP(tsp, control = list(repetitions = 1e3))
ordered_cols <- my_palette[sol]

lab <- convertColor(t(rgb), 'sRGB', 'Lab')
ordered_cols2 <- my_palette[order(lab[, 'L'])]

# Plot ordered colors
my_palette_ordered <- ggplot2::qplot(x = 1:50, y = 1, fill = I(ordered_cols2), geom = 'col', width = 1) + ggplot2::theme_void()

# Add different color palette, theme, etc.
ext_by_month_plot <- ext_by_month_plot + scale_color_gradientn(colors = ordered_cols2)
```

###### Change theme, add title, change axis
```{r}
ext_by_month_plot + theme_light() +
  labs(title = "Monthly Change in Sea Ice Extent, by Year", x = "Month", y = "Extent")
```
