---
title: 'Homework #1: Intro to R'
author: "Mickayla Johnston"
date: "9/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 1. Load quakes with data(quakes). Show what’s there with str() and summary().
```{r}
data("quakes")
```
##### str-ing it:
str(quakes)

##### Summarizing it:
summary(quakes)
```

\

#### 2. Show the entirity of the column "long"
```{r}
quakes[["long"]]
```

\

#### 3. Hey, another useful function - unique()! Apply it to a vector, and you can see what are all of the unique values.
```{r}
unique(quakes[["stations"]])
```

\

#### 4. What unique stations are there? Use length() with unique() to determine how many stations there are.
```{r}
length(unique(quakes[["stations"]]))
```
##### There are 102 stations

\

#### 5.Using range(), what is the range of depths where quakes occur?
```{r}
range(quakes[["depth"]])
```
##### The range is 40 to 680

\

#### Extra Credit: Where was the earthquake of largest magnitude found?  
##### My roundabout way of getting the answer :-) 
```{r}
max(quakes[["mag"]])

match(6.4, quakes[["mag"]])

quakes[152, "stations"]
```
#### The earthquake with the largest magnitude was a 6.4 and was at station 122!  
Note: I'm uploading to github as well ([here](https://github.com/mickayla32/bio607))  
Please let me know if I'm doing this incorrectly :) 
 

